### 2013-11-15 author_list_hps_PAC39.tex
Added author list from PAC39 report.

\newcommand{\red[1]}{{\color{red}{\bf #1}}}
\newcommand{\JLAB}{Thomas Jefferson National Accelerator Facility, Newport News, Virginia 23606}
\newcommand{\CUA}{Catholic University of America, Washington, D.C. 20064}
\newcommand{\OU}{Ohio University,  Athens, Ohio 45701}
\newcommand{\YEREVAN}{Yerevan Physics Institute, 375036 Yerevan, Armenia}
\newcommand{\SCAROLINA}{University of South Carolina, Columbia, South Carolina 29208}
\newcommand{\NSU}{Norfolk State University, Norfolk, Virginia 23504}
\newcommand{\ODU}{Old Dominion University, Norfolk, Virginia 23529}
\newcommand{\genova}{Istituto Nazionale di Fisica Nucleare, Sezione di Genova e Dipartimento di Fisica dell\'Universita, 16146 Genova, Italy}
\newcommand{\SACLAY}{CEA, Centre de Saclay, Irfu/Service de Physique Nucl\'eaire 91191 Gif-sur-Yvette, France}
\newcommand{\ORSAY}{Institut de Physique Nucleaire d'Orsay, IN2P3, BP 1, 91406 Orsay, France}
\newcommand{\ECOLE}{CPhT, Ecole Polytechnique, F 91128 PALAISEAU CEDEX, France}
\newcommand{\UCSC}{University of California, Santa Cruz, CA 95064}
\newcommand{\SUNY}{Stony Brook University, Stony Brook, NY 11794-3800}
\newcommand{\FNAL}{Fermi National Accelerator Laboratory, Batavia, IL 60510-5011}
\newcommand{\HUJ}{Hebrew University of Jerusalem, Jerusalem, Israel}
\newcommand{\UNH}{University of New Hampshire, Department of Physics, Durham, NH 03824}
\newcommand{\PERIMETER}{Perimeter Institute, Ontario, Canada N2L 2Y5}
\newcommand{\RPI}{Rensselaer Polytechnic Institute, Department of Physics, Troy, NY 12181}
\newcommand{\SLAC}{SLAC National Accelerator Laboratory, Menlo Park, CA 94025}
\newcommand{\WNM}{The College of William and Mary, Department of Physics, Williamsburg, VA 23185}
\renewcommand*{\thesection}{\arabic{section}}
\newcommand{\Contact}{Contact person}

\begin{document}

\title{\bf\large{Status of the Heavy Photon Search Experiment at Jefferson Laboratory}}
\author{P. Hansson Adrian, C. Field, N. Graf, M. Graham, G. Haller, R. Herbst, J. Jaros\footnote{Co-spokesperson\label{spoks}}, 
T. Maruyama, J. McCormick, K. Moffeit, T. Nelson, H. Neal, A. Odian, M. Oriunno, S. Uemura, D. Walz}
\affiliation{\SLAC}                                 
 \author{A. Grillo, V. Fadeyev, O. Moreno}
\affiliation{\UCSC}
\author{W. Cooper}
\affiliation{\FNAL}
\author{S. Boyarinov, V. Burkert, A. Deur, H. Egiyan, L. Elouadrhiri, A. Freyberger, F.-X. Girod, 
V. Kubarovsky, Y. Sharabian, S. Stepanyan\footref{spoks}, M. Ungaro, B. Wojtsekhowski}
\affiliation{\JLAB}
\author{R. Essig}
\affiliation{\SUNY}
\author{M. Holtrop\footref{spoks}, K. Slifer, S. K. Phillips}
\affiliation{\UNH} 
\author{A. Fradi, B. Guegan, M. Guidal, S. Niccolai, S. Pisano, E. Rauly, P. Rosier and D. Sokhan}
\affiliation{\ORSAY}
\author{P. Schuster, N. Toro}
\affiliation{\PERIMETER}
\author{N. Dashyan, N. Gevorgyan, R. Paremuzyan, H. Voskanyan}
\affiliation{\YEREVAN}
\author{M. Khandaker, C. Salgado}
\affiliation{\NSU}
\author{M. Battaglieri, R. De Vita}
\affiliation{\genova}
\author{S. Bueltmann, L. Weinstein}
\affiliation{\ODU}
\author{G. Ron}
\affiliation{\HUJ}
\author{P. Stoler, A. Kubarovsky}
\affiliation{\RPI}
\author{K. Griffioen}
\affiliation{\WNM}