### 2014-10-24 author_list_hps_PAC39_compact_alphabetical_order.tex
Fixing Orsay name based on Raphaels input.

\newcommand{\red[1]}{{\color{red}{\bf #1}}}
\newcommand{\JLAB}{Thomas Jefferson National Accelerator Facility, Newport News, Virginia 23606}
\newcommand{\CUA}{Catholic University of America, Washington, D.C. 20064}
\newcommand{\OU}{Ohio University,  Athens, Ohio 45701}
\newcommand{\YEREVAN}{Yerevan Physics Institute, 375036 Yerevan, Armenia}
\newcommand{\SCAROLINA}{University of South Carolina, Columbia, South Carolina 29208}
\newcommand{\NSU}{Norfolk State University, Norfolk, Virginia 23504}
\newcommand{\ODU}{Old Dominion University, Norfolk, Virginia 23529}
\newcommand{\genova}{Istituto Nazionale di Fisica Nucleare, Sezione di Genova e Dipartimento di Fisica dell\'Universita, 16146 Genova, Italy}
\newcommand{\SACLAY}{CEA, Centre de Saclay, Irfu/Service de Physique Nucl\'eaire 91191 Gif-sur-Yvette, France}
\newcommand{\ORSAY}{Institut de Physique Nucl\'eaire, CNRS/IN2P3 and Universit\'e Paris Sud, Orsay, France}
\newcommand{\ECOLE}{CPhT, Ecole Polytechnique, F 91128 PALAISEAU CEDEX, France}
\newcommand{\UCSC}{Santa Cruz Institute for Particle Physics, University of California, Santa Cruz, CA 95064}
\newcommand{\SUNY}{Stony Brook University, Stony Brook, NY 11794-3800}
\newcommand{\FNAL}{Fermi National Accelerator Laboratory, Batavia, IL 60510-5011}
\newcommand{\UNH}{University of New Hampshire, Department of Physics, Durham, NH 03824}
\newcommand{\PERIMETER}{Perimeter Institute, Ontario, Canada N2L 2Y5}
\newcommand{\RPI}{Rensselaer Polytechnic Institute, Department of Physics, Troy, NY 12181}
\newcommand{\SLAC}{SLAC National Accelerator Laboratory, Menlo Park, CA 94025}
\newcommand{\WNM}{The College of William and Mary, Department of Physics, Williamsburg, VA 23185}
\newcommand{\GLASGOW}{University of Glasgow, Glasgow, G12 8QQ, Scotland, UK}
\newcommand{\RUTGERS}{Rutgers University, Department of Physics and Astronomy, Piscataway, NJ 08854}

\author[GENOVA]{M. Battaglieri} 
\author[JLAB]{S. Boyarinov} 
\author[ODU]{S. Bueltmann} 
\author[JLAB]{ V. Burkert} 
\author[GENOVA]{A. Celentano} 
\author[ORSAY]{ G. Charles}
\author[FNAL]{W. Cooper}
\author[JLAB]{C. Cuevas}
\author[YEREVAN]{N. Dashyan} 
\author[GENOVA]{ R. DeVita}
\author[ORSAY]{ C. Desnault}
\author[JLAB]{ A. Deur} 
\author[JLAB]{ H. Egiyan}
 \author[JLAB]{ L. Elouadrhiri} 
\author[SUNY]{R. Essig}
\author[UCSC]{ V. Fadeyev} 
\author[SLAC]{C. Field}
 \author[JLAB]{ A. Freyberger} 
\author[RUTGERS]{Y. Gershtein}
\author[YEREVAN]{ N. Gevorgyan} 
 \author[JLAB]{ F.-X. Girod} 
\author[SLAC]{N. Graf} 
\author[SLAC]{ M. Graham} 
\author[WNM]{K. Griffioen}
\author[UCSC]{A. Grillo} 
\author[ORSAY]{ M. Guidal} 
\author[SLAC]{ G. Haller} 
\author[SLAC]{P. Hansson Adrian\corref{corrauthor}}
\ead{phansson@slac.stanford.edu}
\author[SLAC]{ R. Herbst} 
\author[UNH]{M. Holtrop}
\author[SLAC]{ J. Jaros}
 \author[JLAB]{ S. Kaneta}
\author[NSU]{M. Khandaker} 
\author[RPI]{A. Kubarovsky}
\author[JLAB]{ V. Kubarovsky} 
\author[SLAC]{T. Maruyama} 
\author[SLAC]{ J. McCormick} 
\author[SLAC]{ K. Moffeit} 
\author[UCSC]{ O. Moreno}
\author[SLAC]{ H. Neal} 
\author[SLAC]{ T. Nelson} 
\author[ORSAY]{ S. Niccolai} 
\author[SLAC]{ A. Odian} 
\author[SLAC]{ M. Oriunno} 
\author[YEREVAN]{ R. Paremuzyan}
\author[SLAC]{ R. Partridge} 
\author[UNH]{ S. K. Phillips}
\author[ORSAY]{ E. Rauly}
 \author[JLAB]{ B. Raydo} 
\author[RUTGERS]{J. Reichert}
\author[ORSAY]{ E. Rindel} 
\author[ORSAY]{ P. Rosier}
\author[NSU]{ C. Salgado}
\author[PERIMETER]{P. Schuster} 
\author[JLAB]{ Y. Sharabian} 
\author[GLASGOW]{D. Sokhan}
\author[JLAB]{ S. Stepanyan}
\author[PERIMETER]{ N. Toro}
\author[SLAC]{ S. Uemura} 
\author[JLAB]{ M. Ungaro} 
\author[YEREVAN]{ H. Voskanyan}
\author[SLAC]{ D. Walz}
\author[ODU]{ L. B. Weinstein}
\author[JLAB]{ B. Wojtsekhowski}
%
%
%
%
%
%
 %
%
%
%
%
%
%
%
%

\address[GENOVA]{\genova}
\address[JLAB]{\JLAB}
\address[ODU]{\ODU}
\address[FNAL]{\FNAL}
\address[YEREVAN]{\YEREVAN}
\address[ORSAY]{\ORSAY}
\address[SUNY]{\SUNY}
\address[UCSC]{\UCSC}
\address[SLAC]{\SLAC}                                 
\address[RUTGERS]{\RUTGERS}
\address[WNM]{\WNM}
\address[UNH]{\UNH}
\address[NSU]{\NSU}
\address[RPI]{\RPI}
\address[PERIMETER]{\PERIMETER}
\address[GLASGOW]{\GLASGOW}

\cortext[corrauthor]{Corresponding author.}